A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-06-14 paper/fascd.tex
edits through end of V-cycle section

\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, it follows that
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0 \label{eq:fe:downwardsuminclusion}
\end{equation}
for $j=0,1,\dots,J$.  In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
\end{lemma}

@@

It is tempting to try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$.  Specifically, one might split $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and apply unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately, with $\Pi_i$ applied to $\min\{z^j,0\}$ and a straightforward variant operator to $\max\{z^j,0\}$.  In fact, the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ holds, as shown in the above proof.  However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$, so unfortunately $\Pi_i$ does not map even non-positive functions to $\mathcal{D}^i$ in the presence of a finite upper LDC.

\begin{example}  \label{ex:notfullcd}
Let $a\ge 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Recalling Table \ref{tab:ldcs}, note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other fine-level nodes.  Then $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, for any valid $\overline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$.  Thus, for general $w^1 \in \mathcal{U}^1$ and general LDCs, $\Pi_1$ does not map $\min\{w^1,0\}$ into $\mathcal{D}^1$.
\end{example}

@@

Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.  FIXME this will be used.  While similar to Lemma \ref{lem:downwardadmissibility}, the next Lemma does not follow logically from it.

\begin{lemma}  \label{lem:upwardadmissibility}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
\end{equation}
In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} is a full CD \eqref{eq:constraintdecomp}. \end{lemma}

\begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.  (These facts replace \eqref{eq:fe:telescoping}.)  If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.

In unilateral cases we follow the proof of Lemma \ref{lem:downwardadmissibility}.  Definition \eqref{eq:fe:minimummaps} is unchanged.  We can use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but it must be modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.  The remainder of the proof goes through as for Lemma \ref{lem:downwardadmissibility}.
\end{proof}

In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:fascd}.  Lemma \ref{lem:downwardadmissibility} explains downward admissibility, and Lemma \ref{lem:upwardadmissibility} upward admissibility, in Algorithm \ref{alg:fascd}.  However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas.

@@

The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  \fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.  Note that we coarsen both the residual and the solution iterate, in constrast to classical linear multigrid wherein only the residual is coarsened.

However, the corrected iterate needs to be admissible.  In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:implementation}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.  Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.  Then the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.

@@

A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.  Intuitively-speaking, going downward one does not know what the upcoming coarse corrections will be, so they must be in the sufficiently-small sets $\mathcal{D}^j$ so that their sum is admissible once we return to a level.  However, going upward no further \emph{coarser} corrections, relative to that level, can violate admissibility.

To state \fascd Algorithm \ref{alg:fascd} below we must clarify the problems solved at each mesh level.  As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.  On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature and etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.  Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$, and an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{Conceptually-speaking this is true.  In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.  Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}).

As we descend $j<J$, we define new $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
\end{equation}
the value \emph{before} smoothing and/or correction at level $j$.  From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens linear functionals.  (Recall Table \ref{tab:transfers} summarizes our transfer operators.)

On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set, for $z^j \in \mathcal{U}^j$:
\begin{equation}
\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) correction on the level,\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in FE representation, from $\mathcal{V}^{j-1}$ to $\mathcal{V}^{j}$.}
\begin{equation}
z_{(0)}^j = y^j + P z^{j-1},  \label{eq:fe:upwardinitial}
\end{equation}
but with $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.

If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.  In this sense $w^j$ is admissible with respect to the box constraints, but, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate such box constraints $\underline{\gamma}^j$ and $\overline{\gamma}^j$ on each level.  (Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:fascd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.

Regarding formula \eqref{eq:fe:levelsource}, the formula for source functional $\ell^j$ is a classical FAS construction.  (Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE version, which simultaneously justifies \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.  First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level so that
\begin{equation}
\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
\end{equation}
for all $v\in \mathcal{D}^j$.  However, $y$ will actually be approximated from solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
\begin{enumerate}
\item subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
\item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
\item and, where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
\end{enumerate}
Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth as a result of the progress made by the $j+1$ level down-smoother \cite[subsection 5.3.4, for example]{Trottenbergetal2001}.  These steps now yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j} \notag
\end{align}
This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.  The explanation for VI \eqref{eq:fe:upvi} is essentially identical.

We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).  Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved quite accurately.  However, this coarsest-level solver may be the same as the other smoothers, perhaps with a stronger tolerance.

These ideas come together in Algorithm \ref{alg:fascd}.  \pr{fascd-vcycle} acts in-place on $w^J$, and it assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures.  These do in-place modifications of their final arguments, without modifying their first four arguments.  $\pr{smooth}^{\id{down}}$ and $\pr{smooth}^{\id{up}}$ denote \id{down} and \id{up} iterations, respectively.

>>>>

\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, it follows that
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0 \label{eq:fe:downwardsuminclusion}
\end{equation}
for $j=0,1,\dots,J$.  In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
\end{lemma}

@@

One might try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.  In fact, the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ holds, as shown in the above proof.  However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.  In fact, $\Pi_i$ does not map all non-positive functions into $\mathcal{D}^i$ in the presence of a finite upper LDC.

\begin{example}  \label{ex:notfullcd}
Let $a\ge 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Recalling Table \ref{tab:ldcs}, note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.  Then, for any valid $\overline{\chi}^1$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
\end{example}

@@

Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary coarser level.  The incomplete CD in the next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, will justify upward admissibility in the V-cycle in the next Section.

\begin{lemma}  \label{lem:upwardadmissibility}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
\begin{equation}
\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
\end{equation}
In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} becomes a full CD. \end{lemma}

\begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.  If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.

In unilateral cases we follow the proof of Lemma \ref{lem:downwardadmissibility}.  We can use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but it must be modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.  The remainder of the proof goes through as before.
\end{proof}

In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in the V-cycle algorithm which follows.  However, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas, and we cannot offer that analysis.

@@

The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation storage (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

Our V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Note we use prolongation $P$ to indicate where the computer representation changes.} but the corrected iterate needs to be admissible.  Suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:implementation}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.  Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

@@

A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.  Intuitively-speaking, going downward one does not know what the next coarse corrections will be, thus they must be in small sets $\mathcal{D}^j$, so that their sum is admissible once we return to a given level.  On the other hand, going upward no further coarse corrections can violate admissibility.

To state the V-cycle algorithm below we must also clarify the problems solved at each mesh level.  As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.  On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature etc.~by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.  Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$.  An admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.} been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); these LDCs do not depend on the coarse-level corrections.

As we descend we define new $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
\end{equation}
(These are the states \emph{before} smoothing and/or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens linear functionals.\footnote{Recall that Table \ref{tab:transfers} summarizes our transfer operators.}

On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$, in a larger admissible set:
\begin{equation}
\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:\footnote{Prolongation $P$ in \eqref{eq:fe:upwardinitial} reflects the change in FE representation.}
\begin{equation}
z_{(0)}^j = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
\end{equation}
An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.

If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.  In this sense $w^j$ is admissible with respect to the box constraints.  However, while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$, and furthermore, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate these $\underline{\gamma}^j, \overline{\gamma}^j$.  In any case, Algorithm \ref{alg:fascd} yields a new finest-level iterate $w^J+z^J \in \cK^J$ which is  admissible for problem \eqref{eq:fe:vi}.

Observe that formula \eqref{eq:fe:levelsource} for functional $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.  It has the following explanation, mimicking the PDE version, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.  First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
\begin{equation}
\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
\end{equation}
for all $v\in \mathcal{D}^j$.  However, in fact $y$ will be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
\begin{enumerate}
\item subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
\item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction, and
\item where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
\end{enumerate}
Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth; such smoothness should result from application of the down-smoother.  These steps yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
\begin{align}
&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j} \notag
\end{align}
This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} create \eqref{eq:fe:downvi}.  The explanation for VI \eqref{eq:fe:upvi} is similar.

We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).  Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved accurately.

These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which acts in-place on $w^J$.  It assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures; they do in-place modifications of their final arguments, without modifying their first four arguments.  Parameters \id{down} and \id{up} determine the number of smoother iterations, respectively.

\begin{pseudofloat}[ht]
\begin{pseudo}
\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
    $\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
      $\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
      $\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
    $w^J \gets w^J+z^J$
\end{pseudo}
\caption{The \fascd V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
\label{alg:fascd}
\end{pseudofloat}
